Fast approximate Bayesian inference for small-area estimation of HIV indicators using the Naomi model

### 2023-03-10 src/docs_bayescomp-poster/bayescomp-poster.Rmd
Final changes because headless_chrome is different to headless_firefox and my life could have been so much easier making this in PowerPoint

<!-- 38761C -->

@@

@@

* Latent Gaussian models (LGMs) [@rue2009approximate] are three stage hierarchical models with observations $y$, Gaussian latent field $x$ and hyperparameters $\theta$
* In an LGM the conditional mean depends on exactly one structured additive predictor $\mu_i = g(\eta_i)$ with $g: \mathbb{R} \to \mathbb{R}$
  * The `R-INLA` implementation of integrated nested Laplace approximations applies only to LGMs, because ELGM precision matrices are not as sparse
* Extended latent Gaussian models (ELGM) remove this requirement such that $\mu_i = g(\eta_{\mathcal{J_i}})$ where $g_i: \mathbb{R}^{|\mathcal{J_i}|} \to \mathbb{R}$ and $\mathcal{J_i}$ is some set of indices
  * Allows a higher degree of non-linearity in the model
* Naomi is an ELGM, not an LGM, because it includes complex dependency structures:
  1. ANC indicators offset from household survey
  2. Incidence depends on prevalence and ART coverage
  3. Observed data are aggregated finer processes
  4. Allow attendance of ART clinics outside home district
* We extend work of @stringer2022fast in this setting to the challenging Naomi ELGM
* Though we focus on Naomi here, the HIV Inference Group (`hiv-inference.org`) works on many other complex models, challenging for existing Bayesian inference methods, which require high customisability

$$\\[0.4in]$$

@@

* Relatively small country but still a large model: latent field $\text{dim}(x) = 491$, hyperparameters $\text{dim}(\theta) = 24$

$$\\[0.2in]$$

* Fit four inference methods (using one [!] C++ template):
  * `TMB` (`r round(time_taken$TMB, 0)` mins)
  * `aghq` (`r round(time_taken$aghq, 0)` mins): $k = 1$
  * `adam` (`r round(time_taken$adam, 0)` min): $k = 1$
  * `tmbstan` (`r round(time_taken$tmbstan, 1)` days): 4 chains of 100,000 thinned by 40 (required for good diagnostics)
* Figure \@ref(fig:fig3) illustrates example model outputs: HIV prevalence, ART coverage, HIV incidence, and number of new infections, at the district level

>>>>

@@

$$\\[0.1in]$$

@@

* Latent Gaussian models (LGMs) [@rue2009approximate] are three stage hierarchical models with observations $y$, Gaussian latent field $x$ and hyperparameters $\theta$
* In an LGM the conditional mean depends on exactly one structured additive predictor $\mu_i = g(\eta_i)$ with $g: \mathbb{R} \to \mathbb{R}$
  * The `R-INLA` implementation of integrated nested Laplace approximations applies only to LGMs, because ELGM precision matrices are not as sparse
* Extended latent Gaussian models (ELGM) remove this requirement such that $\mu_i = g(\eta_{\mathcal{J_i}})$ where $g_i: \mathbb{R}^{|\mathcal{J_i}|} \to \mathbb{R}$ and $\mathcal{J_i}$ is some set of indices
  * Allows a higher degree of non-linearity in the model
* Naomi is an ELGM, not an LGM, because it includes complex dependency structures:
  1. ANC indicators offset from household survey
  2. Incidence depends on prevalence and ART coverage
  3. Observed data are aggregated finer processes
  4. Allow attendance of ART clinics outside home district
* We extend work of @stringer2022fast in this setting to the challenging Naomi ELGM
* Though we focus on Naomi here, the HIV Inference Group (`hiv-inference.org`) works on many other complex models, challenging for existing Bayesian inference methods, which require flexible modelling tools

@@

* Relatively small country but still a large model: latent field $\text{dim}(x) = 491$, hyperparameters $\text{dim}(\theta) = 24$
* Fit four inference methods (using one [!] C++ template):
  * `TMB` (`r round(time_taken$TMB, 0)` mins)
  * `aghq` (`r round(time_taken$aghq, 0)` mins): $k = 1$
  * `adam` (`r round(time_taken$adam, 0)` min): $k = 1$
  * `tmbstan` (`r round(time_taken$tmbstan, 1)` days): 4 chains of 100,000 thinned by 40 (required for good diagnostics)
* Figure \@ref(fig:fig3) illustrates example model outputs: HIV prevalence, ART coverage, HIV incidence, and number of new infections, at the district level
